---
title: "Clustering of HPLC samples"
output:
  html_document:
    code_folding: "hide"
---

# Exploration and clustering of the Boussole HPLC dataset

This work aims to explore the raw pigment data of the year of concomittant HPLC and 3X1M sampling at the Boussole site. </br>
The ultimate goal is to evaluate the possibility of using the 3X1M fluorometer to get informations on the phytoplankton community composition. </br>
To do so, we will try to create a classification model that will allow to predict the type of phytoplankton community. </br>
The "type of phytoplankton community" must be defined in this notebook. We want to evaluate the major successions in pigment composition and the co-appearance of pigments to create a couple of cluster that occurs at given depth and time in the year. </br>
</br>
***
</br>
At first, we load the HPLC data, that have been already match with 3X1M data.

```{r message=FALSE, warning=FALSE}
library(tidyverse)

local_path <- here::here() #define the local path
data_path <- paste(local_path, "Boussole/Output/Data/Compiled/echo_hplc_bouss.csv", sep ="/")

hplc <- read_csv(data_path)

#define some pigment vec to ease the data manipulation later 
pigments <- c("chl_c1_c2", "chl_c3", "peri", "but", "fuco", "neox", "prasi", "viola", "hex", "diad", "allo", "diat", "zea", "lutein", "dv_chlb", "chlb", "t_chlb", "dv_chla", "chla", "t_chla")
pigtosum <- c("chl_c1_c2", "chl_c3", "peri", "but", "fuco", "neox", "prasi", "viola", "hex", "diad", "allo", "diat", "zea", "lutein", "t_chlb", "t_chla")
pigment_to_plot <- pigments[!pigments %in% c("t_chla", "t_chlb", "chla")]
data_to_plot <- select(hplc, bouss, date, depth, sumpig, pigments) %>% 
  mutate_at(all_of(pigment_to_plot), ~./sumpig) %>%
  pivot_longer(all_of(pigment_to_plot), values_to = "proportion", names_to = "pigment")
```

We will visualize each sample as a dot, the size of the dot is the Chla concentration and the colour correspond to the fraction of the pigment in the total accessories pigment concentration.

```{r fig.align="center", fig.width=10, fig.height=6, fig.cap="Figure: Proportion of each pigment in the total pigment concentration for each HPLC sample at BOUSSOLE"}
g <- ggplot(data_to_plot)+
  geom_point(aes(x = date, y = -depth, colour = proportion, size = t_chla))+
  scale_colour_viridis_c(name = "Pigment proportion")+
  xlab("Date")+
  ylab("Depth (m)")+
  scale_size(name = "[Chla]")+
  facet_wrap(.~pigment)
g
```
We observe some interesting features here. First, the Chla concentration start to increase in March/April and deepen during the following month. This is coherent with what has been previously described. We can notice the early increase of the [Chla] and the magnitude of it, we reach 3 mg.m-3 !! </br>
This increase of Chla is associated with an increase of the fucoxanthin proportion indicating of a bloom of Diatom. After which, the diatoms are mainly present in the deep layer (< 50m). The 19'-HF pigment represent more that 10% of the pigment community in almost every samples. we also notice an outbreak of Zeaxanthin in the surface/subsurface layer in april, after which the zeaxanthin represent a significant part of the pigment composition at the surface. Finally, we can observe a depth community with a significant concentration of 19'-bf and Chlb/Chlc. </br>
</br>
We will now reduce the number of dimension (i.e. the number of descriptors) by performing a CA, that consist of regrouping samples with similar relativ abundances (here concentration). This will allow us to perform hierarchical clustering.</br>
```{r fig.align="center", fig.cap="Figure: CA rpojection of HPLC data", fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
library(vegan)

pigments_afc <- pigments[pigments != "t_chla" & pigments != "t_chlb"]

AFC <- cca(select(hplc, all_of(pigments_afc)))

scores <- data.frame(scores(AFC, choices = c(1,2,3), display = "site"))
data_ca <- bind_cols(hplc, scores)

pigscore <- data.frame(scores(AFC, choices = c(1,2,3), display = "species"))


ggplot(data_ca)+
  geom_point(aes(x = CA1, y = CA2, colour = depth, size = t_chla))+
  geom_segment(aes(x = 0, xend = CA1, y = 0, yend = CA2), data = pigscore)+
  ggrepel::geom_text_repel(aes(x = CA1, y = CA2, label = rownames(pigscore)), data = pigscore)+
  scale_color_viridis_c(direction = -1)+
  xlim(-2,2)+
  ylim(-2,2)+
  theme_bw()
```
This analysis already separate samples of different depth, chla concentration... Let's see what clustering can give us.</br>
```{r fig.align="center", fig.cap="Figure: Clustering tree", fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
data_ca <- data_ca %>% mutate("f532_f440" = fluo_532 /fluo_440)
distbouss <- dist(select(data_ca, f440_f470, f532_f470, CA1, CA2))

plot(hclust(distbouss, method = "ward.D"))

data_ca$group <- as.factor(cutree(hclust(distbouss, method = "ward.D"),  h = 20))
```
Following this tree I propose to cut at h = 20.

```{r message=FALSE, warning=FALSE}
data_to_plot <- select(data_ca, bouss, date, depth, sumpig, pigments, group) %>% 
  mutate_at(all_of(pigment_to_plot), ~./sumpig) %>%
  pivot_longer(all_of(pigment_to_plot), values_to = "proportion", names_to = "pigment")
```

```{r fig.align="center", fig.cap="Figure: cluster composition", fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
# cluster visualization ---------------------------------------------------
library(treemap)
library(treemapify)
library(patchwork)

cluster_viz <- data_ca %>% select(pigment_to_plot, group) %>%  group_by(group) %>%
  summarise_all(mean, na.rm = "TRUE") %>% ungroup() %>% 
  pivot_longer(all_of(pigment_to_plot), values_to = "concentration", names_to = "pigment")

tplot <- data_ca %>% 
  group_by(group) %>% 
  mutate(wdp = 1.56 * fuco + 0.92 * peri + 4.8 * allo + 1.02 * but + 1.12 * hex + 1.51 * zea + 0.69 * t_chlb,
         micro = (1.56 * fuco + 0.92 * peri)/wdp,
         nano = (4.8 * allo + 1.02 * but + 1.51 * hex)/wdp,
         pico = (1.51 * zea + 0.69 * t_chlb)/wdp) %>% 
  summarise_at(vars(c(pico, nano, micro, t_chlb, fuco, zea, peri, allo, hex, but, dv_chla)), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  pivot_longer(t_chlb:dv_chla, names_to = 'pigment', values_to = 'concentration') %>% 
  mutate(size = ifelse(pigment %in% c('zea', 't_chlb', 'dv_chla'), 'pico', ifelse(pigment %in% c('allo', 'hex', 'but'), 'nano', ifelse(pigment %in% c('fuco', 'peri'), 'micro', 'error'))))

tplot1 <- filter(tplot, group == '1')
tplot2 <- filter(tplot, group == '2')
tplot3 <- filter(tplot, group == '3')
#tplot4 <- filter(tplot, group == '4')



ga <- ggplot(data_to_plot)+
  geom_point(aes(x = date, y = -depth, size = t_chla), colour = "Grey")+
  geom_point(data = filter(data_to_plot, group == "1"), aes(x =date, y = -depth, size = t_chla), colour = "#e41a1c")+
  scale_size(guide = "none")+
  xlab("")+
  theme_bw()+
  ggtitle("Cluster 1")

gb <- ggplot(data_to_plot)+
  geom_point(aes(x = date, y = -depth, size = t_chla), colour = "Grey")+
  geom_point(data = filter(data_to_plot, group == "2"), aes(x =date, y = -depth, size = t_chla), colour = "#377eb8")+
  xlab("")+
  ylab("")+
  theme_bw()+
  scale_size(guide = "none")+
  ggtitle("Cluster 2")

gc <- ggplot(data_to_plot)+
  geom_point(aes(x = date, y = -depth, size = t_chla), colour = "Grey")+
  geom_point(data = filter(data_to_plot, group == "3"), aes(x =date, y = -depth, size = t_chla), colour = "#4daf4a")+
  scale_size(guide = "none")+
  theme_bw()+
  xlab("")+
  ylab("")+
  ggtitle("Cluster 3")

#gd <- ggplot(data_to_plot)+
  # geom_point(aes(x = date, y = -depth, size = t_chla), colour = "Grey")+
  # geom_point(data = filter(data_to_plot, group == "4"), aes(x =date, y = -depth, size = t_chla), colour = "#984ea3")+
  # scale_size(guide = "none")+
  # theme_bw()+
  # xlab("")+
  # ylab("")+
  # ggtitle("Cluster 4")
#create the three treeplot

g2 <- ggplot(tplot1, aes(area = concentration, fill = size, subgroup = size, label = pigment))+
  geom_treemap(layout = 'fixed')+
  geom_treemap_subgroup_text(layout = 'fixed', place = 'middle', fontface = 'bold', size = 14)+
  geom_treemap_text(layout = 'fixed', place = 'bottomright', 'size' = 11, colour = 'white', fontface = 'italic')+
  guides(fill = "none")+
  scale_fill_brewer(palette = 'Dark2')

g3 <- ggplot(tplot2, aes(area = concentration, fill = size, subgroup = size, label = pigment))+
  geom_treemap(layout = 'fixed')+
  geom_treemap_subgroup_text(layout = 'fixed', place = 'middle', fontface = 'bold', size = 14)+
  geom_treemap_text(layout = 'fixed', place = 'bottomright', 'size' = 11, colour = 'white', fontface = 'italic')+
  guides(fill = "none")+
  scale_fill_brewer(palette = 'Dark2')

g4 <- ggplot(tplot3, aes(area = concentration, fill = size, subgroup = size, label = pigment))+
  geom_treemap(layout = 'fixed')+
  geom_treemap_subgroup_text(layout = 'fixed', place = 'middle', fontface = 'bold', size = 14)+
  geom_treemap_text(layout = 'fixed', place = 'bottomright', 'size' = 11, colour = 'white', fontface = 'italic')+
  guides(fill = "none")+
  scale_fill_brewer(palette = 'Dark2')

#g5 <- ggplot(tplot4, aes(area = concentration, fill = size, subgroup = size, label = pigment))+
  # geom_treemap(layout = 'fixed')+
  # geom_treemap_subgroup_text(layout = 'fixed', place = 'middle', fontface = 'bold', size = 14)+
  # geom_treemap_text(layout = 'fixed', place = 'bottomright', 'size' = 11, colour = 'white', fontface = 'italic')+
  # guides(fill = "none")+
  # scale_fill_brewer(palette = 'Dark2')

(ga|gb|gc) /(g2 | g3 | g4)


```
We find the same structure as in the first structure. The first cluster corresponds to the high concentration of [Chlb] and might not be significative here. 

```{r message=FALSE, warning=FALSE}
results_path <- paste(local_path, "Boussole/Data/labo/results_reshape.csv", sep ="/")

results <- read_csv(results_path)

pigments <- c("chl_c1_c2", "chl_c3", "peri", "but", "fuco", "diad", "diat", "zea", "dv_chlb", "chlb", "t_chlb", "dv_chla", "chla", "t_chla")

result_long <- results %>%
  rename("chl_c1_c2" = chlorophyll_c1_c2_mg_dvp,
         "chl_c3" = chlorophyll_c3,
         "peri" = peridinin,
         "but" = x19_butanoyloxyfucoxanthin,
         "fuco" = fucoxanthin,
         "diad" = diadinoxanthin,
         "diat" = diatoxanthin,
         "zea" = zeaxanthin,
         "dv_chlb" = divinyl_chlorophyll_b,
         "chlb" = chlorophyll_b,
         "dv_chla" = divinyl_chlorophyll_a,
         "chla" = chlorophyll_a,
         "t_chla" = total_chlorophyll_a,
         "f440" = "440nm_mean",
         "f470" = "470nm_mean",
         "f532" = "532nm_mean",
         "fv_fm" = "Fv/Fm") %>% 
  filter(!is.na(number_cells)) %>% 
  pivot_longer(pigments, names_to = "pigment", values_to = "concentration") %>% 
  mutate(cell_conc = concentration/number_cells,
         "f440_f470" = f440/f470,
         "f532_f470" = f532/f470) %>% 
  select(strain, replicate, f440, f470, f532, f440_f470, f532_f470, number_cells, fv_fm, pigment, concentration, cell_conc)

pig_cell = paste(result_long$pigment, "_cell", sep ="")

result_wide <- result_long %>% 
  select(-cell_conc) %>% 
  pivot_wider(names_from = pigment, values_from = concentration)
result_cell_conc <-  result_long %>%
  mutate(pig_cell = paste(pigment, "_cell", sep = "")) %>%
  select(-concentration, -pigment) %>% 
  pivot_wider(names_from = pig_cell, values_from = cell_conc)
  
result_tot <- left_join(result_wide, result_cell_conc) %>% mutate(common_name = case_when(
  strain %in% c("RCC1717", "RCC76", "RCC4213") ~ "Diatom",
  strain %in% c("RCC162", "RCC156", "PCC9511") ~ "Prochlorococcus",
  strain %in% c("RCC2379", "RCC2374", "RCC2319") ~ "Synechococcus",
  strain %in% c("RCC100") ~ "Pelagophyte",
  strain %in% c("RCC3006") ~ "Dino"
))

```






